### 2021-04-13 doc/introduction.Rmd
adding the builded vignettes to the repo

---
title: "Spatial Fuzzy CMean with R"
author: "Jeremy Gelb"
date: "02/04/2021"
output:
  rmarkdown::html_vignette:
    fig_width: 5
    fig_height: 5
    toc: true
    toc_depth: 2
    df_print: "tibble"
vignette: >
  %\VignetteIndexEntry{Introduction}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
bibliography: geocmeans_references.bibtex
link-citations: yes
colorlinks: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

This document is a short introduction to the package **geocmeans**. It implements a fuzzy classification method bringing spatial information and neighbouring in its calculation

@cai2007fast described in their article the method, originally applied in the analysis of brain imagery. The generalized version of the spatial fuzzy c-means is presented by @zhao2013kernel.

@gelb2021apport applied the method to socio-residential and environmental data and compared the results with oter unsupervised classification algorithms (in french).

There are actually numerous packages and functions to perform unsupervised classification in R (*hclust*, *kmeans*, *cmeans*, *factoextra*, etc.). However theses methods are not always well-suited to analyze spatial data. Indeed, they do not take into account spatial information like proximity or contiguity between observations. This may lead to solutions for which close observations end up in different groups despite the fact that they are very similar.

The package **ClustGeo** is to our knowledge the only package proposing an unsupervised classification method considering directly spatial proximity between observations. The proposed approach is very appealing because the user has to select a parameter (*alpha*) that controls the weight of the spatial distance matrix (calculated between observations with their locations) VS the semantic distance matrix (calculated  between observations with their variables).

However, this method belongs to the category of "hard-clustering" algorithms. Each observation ends up in one cluster/group. The main draw-back here is the difficulty to identify observations that are undecided, at the frontier of two clusters/groups. The soft or fuzzy clustering algorithms give more information because they calculate for each observation its "probability" of belonging to each group.

The algorithms SFCM (spatial fuzzy c-means) and SGFCM (spatial generalized fuzzy c-means) propose to combine the best of both worlds.

The package **geocmeans** is an implementation in R of these methods (originally developed for analysis of brain imagery). It comes with a set of functions to facilitate the analysis of the final membership matrices :

* calculating many quality indices (coming mainly from the package *fclust*)
* mapping the results
* giving summary statistics for each cluster/group.

## Loading the packages and the data

**geocmeans** comes with a toy dataset *LyonIris*, combining many demographic and environmental variables aggregated at the scale of the Iris (aggregated units for statistical information) in Lyon (France).

Before starting the analysis, it is required to standardize the data because most of the calculus is based on Euclidean distance (we plan to also include Manhattan distance in a future release).

```{r message=FALSE, warning=FALSE}
#charging packages and data
library(geocmeans)
library(ggplot2)
library(ggpubr)
library(dplyr)
library(viridis)
library(spdep)

data(LyonIris)

#selecting the columns for the analysis
AnalysisFields <-c("Lden","NO2","PM25","VegHautPrt","Pct0_14",
                   "Pct_65","Pct_Img","TxChom1564","Pct_brevet","NivVieMed")

#rescaling the columns
Data <- LyonIris@data[AnalysisFields]
for (Col in names(Data)){
  Data[[Col]] <- scale(Data[[Col]])
}

#preparing some elements for further mapping
LyonIris$OID <- as.character(1:nrow(LyonIris))
FortiData <- broom::tidy(LyonIris,region="OID")
```

```{r include=FALSE}
#loading the pre-calculated results
load(system.file("extdata", "results_vignette_intro.rda",
                           package = "geocmeans", mustWork = TRUE))
```

## Classical Kmeans

To explore the dataset and choose a number of cluster/groups (k) we propose to start with a classical kmeans.

```{r warning=FALSE, fig.cap = "Impact of the number of groups on the explained variance", out.width = "50%", fig.pos="H", fig.align="center"}
#finding the best k by using the r2 of the classification
#trying for k from 2 to 10
R2s <- sapply(2:10,function(k){
  Clust <- kmeans(Data,centers=k,iter.max = 150)
  R2 <- Clust$betweenss / Clust$totss
  return(R2)
})

Df <- data.frame(K=2:10,
                 R2 = R2s)

ggplot(Df)+
  geom_line(aes(x=K,y=R2s))+
  geom_point(aes(x=K,y=R2s),color="red")+
  xlab("Number of groups")+
  ylab("R2 of classification")
```

By plotting the R-squared of the kmeans classification for each k between 2 and 10, we can see a first elbow at k=3. But this small number of groups leads to a classification explaining only 43% of the original data variance. We decide to keep k=4 to have one more group in the analysis.

Let us map the obtained groups.

```{r warning=FALSE, fig.cap="Clusters identified by classical k-means", fig.pos="H", fig.align="center"}
KMeanClust <-  kmeans(Data,centers=4,iter.max = 150)
LyonIris$Cluster <-paste("cluster",KMeanClust$cluster,sep="_")

#mapping the groups
DFmapping <- merge(FortiData,LyonIris,by.x="id",by.y="OID")

ggplot(data=DFmapping)+
  geom_polygon(aes(x=long,y=lat,group=group,fill=Cluster),color=rgb(0,0,0,0))+
  coord_fixed(ratio = 1)+
  scale_fill_manual(name="Cluster",values = c("cluster_1"="palegreen3",
                                 "cluster_2"="firebrick",
                                 "cluster_3"="lightyellow2",
                                 "cluster_4"="steelblue"))+
  theme( axis.title = ggplot2::element_blank(),
    axis.text = ggplot2::element_blank(),
    axis.ticks = ggplot2::element_blank()
    )

```

We can clearly distinguish 4 strong spatial structures, but with some mixing between the clusters.

We could now compare this solution with a classical c-means algorithm.

## Classical c-means and generalized c-means

The classical c-means is a simple method to perform fuzzy unsupervised classification. The package **geocmeans** proposes the function *CMeans*. We set the fuzzyness degree (m) to 1.5.

We do not present here the algorithm, but only the two main formulas.

The first one is used to update the values of the membership matrix at each iteration $u_ik$

$$u_{ik} = \frac{(||x_{k} - v{_i}||^{2}) ^{(-1/(m-1))}}{\sum_{j=1}^c(||x_{k} - v{_j}||^2 )^{(-1/(m-1))}}$$
And the second one to update the centers of the clusters

$$v_{i} = \frac{\sum_{k=1}^N u_{ik}^m(x_{k})}{\sum_{k=1}^N u_{ik}^m}$$

With :

* $x_k$ the values of the observation $k$
* $v_i$ the values of the center of the cluster $i$
* $c$ the number of clusters
* $m$ the fuzzyness index

```{r warning=FALSE}
Cmean <- CMeans(Data,4,1.5,500,standardize = FALSE, seed = 456, tol = 0.00001, verbose = FALSE)
```

We can now use the function *calcqualityIndexes* which combines many indices from the package **fclust** to analyze the quality of the classification. We will use these values later to do comparisons among the different algorithms.

```{r message=FALSE, warning=FALSE}
calcqualityIndexes(Data, Cmean$Belongings, m = 1.5)
```

Note, **geocmeans** proposes also a generalized version of the c-means algorithm. It is known to accelerate convergence and yield less fuzzy results by adjusting the membership matrix at each iteration. It requires an extra parameter *beta* controlling the strength of the modification. The modification affects only the formula updating the membership matrix.

$$u_{ik} = \frac{(||x_{k} - v{_i}||^{2} - \beta_k) ^{(-1/(m-1))}}{\sum_{j=1}^c(||x_{k} - v{_j}||^2 - \beta_k)^{(-1/(m-1))}}$$

with $\beta_k = min(||x_{k} - v||^2)$ and $0 \leq \beta \leq 1$.

To select an appropriate value for this parameter, we will try all the possible values between 0 and 1 with a step of 0.05.

```{r warning=FALSE, eval = FALSE}
beta_values <- select_parameters("GFCM",data = Data, k = 4, m = 1.5,
                                 beta = seq(0,1,0.05), spconsist = F,
                                 tol = 0.00001, seed = 456)
```

```{r warning=FALSE}
knitr::kable(beta_values[c("beta","Silhouette.index","XieBeni.index","Explained.inertia")],
             col.names = c("beta", "silhouette index",
                           "Xie and Beni index", "explained inertia"),digits = 3)
```

Considering the table above, we select beta = 0.7, it keeps a good silhouhette index, and increases Xie and Beni index and explained inertia. Let us compare the results of GFCM and FCM.

```{r warning=FALSE}
GCmean <- GCMeans(Data,k = 4,m = 1.5, beta = 0.7,500,standardize = FALSE, seed=456,
                  tol = 0.00001, verbose = FALSE)
r1 <- calcqualityIndexes(Data,GCmean$Belongings,m=1.5)
r2 <- calcqualityIndexes(Data,Cmean$Belongings,m=1.5)
df <- cbind(unlist(r1), unlist(r2))

knitr::kable(df,
             digits = 3,col.names = c("GFCM", "FCM"))
```
The results indicate that the GFCM gives a solution that is less fuzzy (higher explained inertia and lower partition entropy) but keeps a good silhouette index and an even better Xie and Beni index.

We can now map the two membership matrices and the most likely group for each observation. To do so, we use the function *mapClusters* from **geocmeans**. We propose here to define a threshold of 0.45. If an observation has only values below this probability in a membership matrix, it will be labeled as "undecided" (represented with transparency on the map).

We can compare the maps of the classical c-means and the generalized version.
```{r warning=FALSE, fig.cap = "Probability of belonging to cluster 1", out.width="80%", fig.pos="H", fig.align="center"}
cmeansMaps<- mapClusters(LyonIris,Cmean$Belongings,undecided = 0.45)
GcmeansMaps<- mapClusters(LyonIris,GCmean$Belongings,undecided = 0.45)

ggarrange(cmeansMaps$ProbaMaps[[1]],GcmeansMaps$ProbaMaps[[1]], 
          nrow = 1, ncol = 2, common.legend = TRUE, legend = "bottom")
```
```{r warning=FALSE, fig.cap = "Probability of belonging to cluster 2", out.width="80%", fig.pos="H", fig.align="center"}
ggarrange(cmeansMaps$ProbaMaps[[2]],GcmeansMaps$ProbaMaps[[2]], 
          nrow = 1, ncol = 2, common.legend = TRUE, legend = "bottom")
```
```{r warning=FALSE, fig.cap = "Probability of belonging to cluster 3", out.width="80%", fig.pos="H", fig.align="center"}
ggarrange(cmeansMaps$ProbaMaps[[3]],GcmeansMaps$ProbaMaps[[3]], 
          nrow = 1, ncol = 2, common.legend = TRUE, legend = "bottom")
```
```{r warning=FALSE, fig.cap = "Probability of belonging to cluster 4", out.width="80%", fig.pos="H", fig.align="center"}
ggarrange(cmeansMaps$ProbaMaps[[4]],GcmeansMaps$ProbaMaps[[4]], 
          nrow = 1, ncol = 2, common.legend = TRUE, legend = "bottom")
```
```{r warning=FALSE, fig.cap = "Most likely clusters and undecided units", out.width="80%", fig.pos="H", fig.align="center"}
ggarrange(cmeansMaps$ClusterPlot,GcmeansMaps$ClusterPlot,
          nrow = 1, ncol = 2, common.legend = TRUE, legend = "bottom")
```

As expected, the results are very similar, but the generalized version provides a more clear-cut classification.

## Spatial c-means and generalized c-means

### Selecting *alpha* for SFCM

Now we can use the *SFCM* function to perform a spatial c-means. The first step is to define a spatial weight matrix indicating which observations are neighbours and the strength of their relationship. We propose here to use a basic queen neighbour matrix (built with **spdep**). The matrix must be row-standardized to ensure that the interpretation of all the parameters remains clear.

```{r message=FALSE, warning=FALSE}
library(spdep)

Neighbours <- poly2nb(LyonIris,queen = TRUE)
WMat <- nb2listw(Neighbours,style="W",zero.policy = TRUE)
```

The main challenge with the SFCM method is to select the parameter *alpha*. It represents the weight of the spatial dimension (lagged values) in the calculus of the membership matrix and the cluster centers.

* If alpha=0, then we end up with a classical c-means algorithm.
* If alpha=1, then the original and the lagged values have the same weight
* If alpha=2, then the lagged values are twice more important than the original values
* end so on...

The two following formulas show how the functions updating the membership matrix and the centers of the clusters are modified.

$$u_{ik} = \frac{(||x_{k} - v{_i}||^2 + \alpha||\bar{x_{k}} - v{_i}||^2)^{(-1/(m-1))}}{\sum_{j=1}^c(||x_{k} - v{_j}||^2 + \alpha||\bar{x_{k}} - v{_j}||^2)^{(-1/(m-1))}}$$

$$v_{i} = \frac{\sum_{k=1}^N u_{ik}^m(x_{k} + \alpha\bar{x_{k}})}{(1 + \alpha)\sum_{k=1}^N u_{ik}^m}$$
with $\bar{x}$ the lagged version of x and $\alpha \geq 0$

As the formula suggests, the *SFCM* can be see as a spatially smoothed version of the classical c-means and *alpha* controls the degree of spatial smoothness. This smoothing can be interpreted as an attempt to reduce spatial overfitting of the classical c-means.

To select *alpha*, we propose to check all the possible values between 0 and 2 with a step of 0.05.

```{r warning=FALSE,  message=FALSE, eval = FALSE}
DFindices_SFCM <- select_parameters(algo = "SFCM", data = Data,
                               k = 4, m = 1.5, alpha = seq(0,2,0.05),
                               nblistw = WMat, standardize = FALSE,
                               tol = 0.0001, verbose = FALSE, seed = 456)

```

Now we can check the indices to select the best *alpha*. The goal is to reduce as much as possible spatial inconsistency and to keep a good classification quality.

Let's start with the spatial inconsistency. This indicator (developed for this package) calculates the sum of the squared differences between each observation and its neighbours on the membership matrix. Thus, the maximum for each observation is $k*j$ with *j* the number of neighbours for the observation and *k* the number of groups. A maximum is reached if each observations have 100% chance to belong to a cluster that is different from all its neighbours. So, when we sum up the values obtained for all the observations, we obtain a quantity of spatial inconsistency. This quantity is divided by the quantity obtained when permuting randomly the rows of the membership matrix. This second quantity represents the spatial inconsistency that we might expect if the observations were randomly scattered in space. We can repeat the permutation step (Monte Carlo approach) and keep the mean of the ratios to have a more robust indicators (see help(spConsistency) for details).

A smaller value indicates a smaller spatial inconsistency and thus a greater spatial consistency. 0 meaning that all observations have exactly the same values in the membership matrix than their neighbours (perfect spatial consistency).

```{r warning=FALSE, fig.cap = "Link between alpha and spatial inconsistency", out.width="50%", fig.pos="H", fig.align="center"}
ggplot(DFindices_SFCM)+
  geom_smooth(aes(x=alpha,y=spConsistency), color = "black")+
  geom_point(aes(x=alpha,y=spConsistency), color = "red")

```

Unsurprisingly, increasing *alpha* leads to a decrease of the spatial inconsistency. This gain follow an inverse function.

Let's check now the explained inertia
```{r warning=FALSE, fig.cap = "Link between alpha and explained inertia", out.width="50%", fig.pos="H", fig.align="center"}
ggplot(DFindices_SFCM)+
  geom_smooth(aes(x=alpha,y=Explained.inertia), color = "black")+
  geom_point(aes(x=alpha,y=Explained.inertia), color = "red")
```

As expected, the explained inertia decreases when alpha increases and follows again an inverse function. The classification has to find a compromise between the original values and the lagged values. However, the loss is very small here: only 3% between alpha = 0 and alpha = 2.

To finish here, we can observe the silhouette and Xie and Beni indicators.

```{r fig.cap="Link between alpha and silhouette index", message=FALSE, warning=FALSE, fig.pos="H", fig.align="center", out.width="50%"}
ggplot(DFindices_SFCM)+
  geom_smooth(aes(x=alpha,y=Silhouette.index), color = "black")+
  geom_point(aes(x=alpha,y=Silhouette.index), color = "red")
```

```{r fig.cap="Link between alpha and Xie and Beni index", message=FALSE, warning=FALSE, out.width="50%", fig.pos="H", fig.align="center"}
ggplot(DFindices_SFCM)+
  geom_smooth(aes(x=alpha,y=XieBeni.index), color = "black")+
  geom_point(aes(x=alpha,y=XieBeni.index), color = "red")

```

The detail of the meaning of theses indicators is beyond the scope of this vignette. Let's just stress that a bigger silhouette index indicates a better classification, and a smaller Xie and beni index indicates a better classification.

After considering all the previous charts, we decide to keep alpha = 0.7 as it seems to give a good balance between spatial consistency and classification quality in this case.

```{r warning=FALSE}
SFCM <- SFCMeans(Data, WMat, k = 4, m = 1.5, alpha = 0.7,
                 tol = 0.0001, standardize = FALSE,
                 verbose = FALSE, seed = 456)
```

### Selecting alpha and beta for SGFCM

It is also possible to use the generalized version of the spatial c-means. In that case, we have to define both *alpha* and *beta*.

The next formula shows how the membership matrix is updated at each iteration. Note that the centers of the clusters are updated with the same formula as SFCM.

$$u_{ik} = \frac{(||x_{k} - v{_i}||^2 -\beta_k + \alpha||\bar{x_{k}} - v{_i}||^2)^{(-1/(m-1))}}{\sum_{j=1}^c(||x_{k} - v{_j}||^2 -\beta_k + \alpha||\bar{x_{k}} - v{_j}||^2)^{(-1/(m-1))}}$$

Because we select a high resolution for our grid search of *alpha* and *beta*, we will use a multiprocessing approach.

```{r warning=FALSE, eval = FALSE}
future::plan(future::multiprocess(workers=4))
DFindices_SFGCM <- select_parameters.mc(algo = "SGFCM", data = Data,
                               k = 4, m = 1.5, alpha = seq(0,2,0.05),
                               beta = seq(0,0.85,0.05),
                               nblistw = WMat, standardize = FALSE, chunk_size = 50,
                               tol = 0.0001, verbose = FALSE, seed = 456)
```

```{r warning=FALSE, fig.cap = "Impact of beta and alpha on silhouette index", out.width = "80%", fig.pos="H", fig.align="center"}
ggplot(DFindices_SFGCM) + 
  geom_raster(aes(x = alpha, y = beta, fill = Silhouette.index),  size = 5) + 
  scale_fill_viridis() +
  coord_fixed(ratio=1)
```
```{r warning=FALSE, fig.cap = "Impact of beta and alpha on Xie and Beni index", out.width = "80%", fig.pos="H", fig.align="center"}
ggplot(DFindices_SFGCM) + 
  geom_raster(aes(x = alpha, y = beta, fill = XieBeni.index),  size = 5) + 
  scale_fill_viridis() +
  coord_fixed(ratio=1)
```
```{r warning=FALSE, fig.cap = "Impact of beta and alpha on spatial inconsistency", out.width = "80%", fig.pos="H", fig.align="center"}
ggplot(DFindices_SFGCM) + 
  geom_raster(aes(x = alpha, y = beta, fill = spConsistency),  size = 5) + 
  scale_fill_viridis() +
  coord_fixed(ratio=1)

```

The first two plots indicate that some particular pairs of *alpha* and *beta* yield good results in the range 0.8 < *alpha* < 1.2 and 0.6 < *beta* < 0.8. The last plot shows that the selection of *beta* has no impact on the spatial consistency.

Considering the previous plots, we decide to retain the solution with *beta* = 0.65 and *alpha* = 0.95 which yield very good results for all the considered indices.

```{r warning=FALSE}
SGFCM <- SGFCMeans(Data,WMat,k = 4,m=1.5, alpha=0.95, beta = 0.65,
                    tol=0.0001, standardize = FALSE, verbose = FALSE, seed = 456)
```

Again, we compare here the generalized and the classical version of the spatial c-means algorithm.

```{r warning=FALSE}
r1 <- calcqualityIndexes(Data, SFCM$Belongings,m = 1.5)
r2 <- calcqualityIndexes(Data, SGFCM$Belongings,m = 1.5)

diagSFCM <- spatialDiag(belongmatrix = SFCM$Belongings, nblistw = WMat,
                        undecided = 0.45,nrep = 500)
diagSGFCM <- spatialDiag(belongmatrix = SGFCM$Belongings, nblistw = WMat,
                         undecided = 0.45,nrep = 500)

df <- cbind(
  c(unlist(r1),diagSFCM$SpConsist),
  c(unlist(r2),diagSGFCM$SpConsist)
)
row.names(df)[length(row.names(df))] <- "sp.consistency"

knitr::kable(df,digits = 3,col.names = c("SFCM","SGFCM"))
```

The solution of the SGFCM is better on the semantic and the spatial aspects.

We can compare the maps
```{r warning=FALSE, fig.cap = "Probability of belonging to cluster 1", out.width="80%", fig.pos="H", fig.align="center"}
SFCMMaps <- mapClusters(geodata = LyonIris,belongmatrix = SFCM$Belongings,undecided = 0.45)
SGFCMMaps <- mapClusters(geodata = LyonIris,belongmatrix = SGFCM$Belongings,undecided = 0.45)

ggarrange(SFCMMaps$ProbaMaps[[1]],SGFCMMaps$ProbaMaps[[1]], nrow = 1, ncol = 2,
          common.legend = TRUE, legend = "bottom")
```
```{r warning=FALSE, fig.cap = "Probability of belonging to cluster 2", out.width="80%", fig.pos="H", fig.align="center"}
ggarrange(SFCMMaps$ProbaMaps[[2]],SGFCMMaps$ProbaMaps[[2]], nrow = 1, ncol = 2,
          common.legend = TRUE, legend = "bottom")
```
```{r warning=FALSE, fig.cap = "Probability of belonging to cluster 3", out.width="80%", fig.pos="H", fig.align="center"}
ggarrange(SFCMMaps$ProbaMaps[[3]],SGFCMMaps$ProbaMaps[[3]], nrow = 1, ncol = 2,
          common.legend = TRUE, legend = "bottom")
```
```{r warning=FALSE, fig.cap = "Probability of belonging to cluster 4", out.width="80%"}
ggarrange(SFCMMaps$ProbaMaps[[4]],SGFCMMaps$ProbaMaps[[4]], nrow = 1, ncol = 2,
          common.legend = TRUE, legend = "bottom")
```
```{r warning=FALSE, fig.cap = "Most likely cluster and undecided units", out.width="80%", fig.pos="H", fig.align="center"}
ggarrange(SFCMMaps$ClusterPlot,SGFCMMaps$ClusterPlot, nrow = 1, ncol = 2,
          common.legend = TRUE, legend = "bottom")
```

## Comparing the spatial consistency of FCM, GFCM, SFCM and SGFCM

Now, we can  do a deeper spatial analysis and compare the spatial consistency of the four classifications realized (FCM, GFCM, SFCM, SGFCM).

```{r warning=FALSE}
spdiag_1 <- spatialDiag(Cmean$Belongings, nblistw = WMat,nrep=250)
spdiag_2 <- spatialDiag(GCmean$Belongings, nblistw = WMat,nrep=250)
spdiag_3 <- spatialDiag(SFCM$Belongings, nblistw = WMat,nrep=250)
spdiag_4 <- spatialDiag(SGFCM$Belongings, nblistw = WMat,nrep=250)

#looking at the moran I values for each group
moran_table <- data.frame(cbind(spdiag_1$MoranValues$MoranI,
                     spdiag_2$MoranValues$MoranI,
                     spdiag_3$MoranValues$MoranI,
                     spdiag_4$MoranValues$MoranI
                     ))
row.names(moran_table) <- paste("cluster ",1:4,sep="")
knitr::kable(moran_table, digits = 3,
             col.names = c("FCM","GFCM","SFCM","SGFCM"),
             caption = "Moran I index for the columns of the membership matrix"
             )
```

Unsurprisingly, the Moran I values calculated on the membership matrices are higher for SFCM and SGFCM, indicating stronger spatial structures in the classifications.

```{r warning=FALSE}
print(c(spdiag_1$SpConsist, spdiag_2$SpConsist,spdiag_3$SpConsist,spdiag_4$SpConsist))
```

Considering the values of spatial inconsistency, we could check if the value obtained for SGFCM is significantly lower than the one of SFCM. Considering the previous `r length(spdiag_3$SpConsistSamples)` permutated values, we can calculate a pseudo p-value:

```{r warning=FALSE}
sum(spdiag_4$SpConsist > spdiag_3$SpConsistSamples) / length(spdiag_3$SpConsistSamples)
```

It appears that out of `r length(spdiag_3$SpConsistSamples)` permutations, the observed values of spatial inconsistency of SGFCM is always lower than the one of SFCM. The difference is significant at the threshold `r 1/length(spdiag_3$SpConsistSamples)`=1/`r length(spdiag_3$SpConsistSamples)`

We can map the undecided observations of the final solution. These entities should be analyzed more precisely. Selecting them is easy with the function *undecidedUnits* of the **geocmeans** package.

```{r warning=FALSE}
Undecided <- undecidedUnits(SGFCM$Belongings,0.45)
LyonIris$FinalCluster <- ifelse(Undecided=="Undecided",
                                "Undecided",paste("cluster",Undecided,sep="_"))

#mapping the groups
DFmapping <- merge(FortiData,LyonIris,by.x="id",by.y="OID")

ggplot(data=DFmapping)+
  geom_polygon(aes(x=long,y=lat,group=group,fill=FinalCluster),color=rgb(0,0,0,0))+
  coord_fixed(ratio = 1)+
  scale_fill_manual(name="FinalCluster",values = c("cluster_V1"="palegreen3",
                                 "cluster_V2"="firebrick",
                                 "cluster_V3"="lightyellow2",
                                 "cluster_V4"="steelblue",
                                 "cluster_V5"="pink",
                                 "Undecided"=rgb(0,0,0,0.4)))+
  theme( axis.title = element_blank(),
    axis.text = element_blank(),
    axis.ticks = element_blank()
    )
```

## Interpreting the results of the final classification

One can obtain many descriptive information about the final groups with three functions of **geocmeans** :

* *summarizeClusters* : calculate summary statistics for each group for a given dataset by using the membership matrix as weights (sticking with the fuzzy spirit).
* *spiderPlots* : display a spider plot allowing to compare quickly the differences between groups.
* *violinPlots* : display a violin plot for each variable in a given dataset. Observations must be grouped before.

```{r warning=FALSE}
summarizeClusters(LyonIris@data[AnalysisFields],belongmatrix = SGFCM$Belongings,
                  weighted = T,dec = 3)
```

```{r warning=FALSE, eval = FALSE}
spiderPlots(LyonIris@data[AnalysisFields], SGFCM$Belongings,
            chartcolors = c("darkorange3","grey4","darkgreen","royalblue"))
violinPlots(LyonIris@data[AnalysisFields], SGFCM$Groups)
```

That's all folks ! These are the enhancements for the next version

* introduce other methods of spatial c-means
* open some other parameters to the user (like the function defining the convergence criterion)
* work on documentation
* improve calculus speed by dropping some apply in the code
* adding the manhattan distance (useful when data have a high dimensionality)